Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs

### 2014-07-17 omrp_lb_rev.tex
More corrections

The Overlapping Batch Means (OBM) method, proposed by \citet{Meketon1984}, is commonly used in simulation to obtain variance estimators having variances that are comparatively lower than their nonoverlapped counterparts. 
In the context of steady-state simulation, OBM is often used to estimate the variance of the sample mean, which is itself an estimator for the mean. 
\cite{SAH90} use overlapping estimators to estimate the variance of estimators for non-means. 
More recently, overlapping estimators based on standardized time series (as opposed to batch means) have been evaluated \citep{Alexopoulos01012007,Alexopoulos2007}; see also \citep{Meterelliyoz_etal_12}.
To the best of our knowledge, OBM has never before been studied in a stochastic optimization context. 
In this paper, we show that the same benefits of nearly the same bias and lower variances relative to nonoverlapped counterparts can be obtained from overlapping batches when used in a stochastic optimization setting.

@@

The optimality gap $\e{f(\xh,\xit)} - \zs$ often cannot be calculated exactly. First, for a given candidate solution $\xh \in X$, evaluation of $\e{f(\xh,\xit)}$ typically involves a difficult multidimensional integral. 
Second, we do not know $\zs$. 
In many fields of optimization, the second problem is alleviated by obtaining lower bounds on $\zs$ through relaxations such as integrality, Lagrangian, or semidefinite relaxations.
As a result, it is common to evaluate an upper bound on the optimality gap. 
This motivates estimating the upper bound on the optimality gap in stochastic programming by using Monte Carlo simulation.
Given a sample size $m$, an upper bound on the optimality gap can be obtained by $$
\e{f(\xh,\xit)} - \zs \leq \e{f(\xh,\xit)} - \e{\zs_m}
$$ 
due to the inequality $\e{\zs_m} \leq \zs$ \citep{Mak1999,norkin_pflug_ruszczynski_98}.
This bound improves as the sample size increases, that is, $\e{\zs_m} \leq \e{\zs_{m+1}} \leq \zs$.  
A straightforward estimate of $\e{f(\xh,\xit)}$ is the sample mean, $\frac{1}{m} \sum_{i=1}^m f(\xh,\xiti)$. 
Instead of $\e{\zs_m}$, we can simply use $\zs_m$.  
The resulting point estimator of the optimality gap is then 
$$
\frac{1}{m} \sum_{i=1}^m f(\xh,\xiti) - \zs_m.
$$ 
Here, we assume the same observations $\xit^1, \dots, \xit^m$ are used in both terms.

Computing the above optimality gap estimate involves solving an optimization problem (\ref{eq:sto_prog_m}) to obtain a lower bound estimator of $\zs$, $\zs_m$, which complicates the statistical analysis. 
As mentioned above, to enable statistical inference, the MRP of \citet{Mak1999} generates $k$ independent estimators, each using sample size $m$ ($k$ nonoverlapping batches of size $m$) and averages them to obtain a point estimator.  
The sample variance of these estimators is used to form confidence intervals (CIs) for the optimality gap (we further review MRP in \S \ref{ssec:mrp}).  
An advantage of MRP is its applicability to a wide range of problems.  
With i.i.d.\ sampling, $f$ can be linear or nonlinear, $\X$ can include integrality constraints or not.  
It is also easy to implement; thus, MRP has been applied to a wide variety of problems, including finance, supply chain network design, etc.; see, e.g., \citep{bertocchi_etal_99,janjarassuk_linderoth_08,santoso_ahmed_etal_05}.  
Recently, the approach of (nonoverlapping) batching has been used for assessing solution quality of stochastic programs with finitely many expected value \citep{wang_ahmed_08} and stochastic dominance \citep{hu2010sample} constraints.

@@

Our aim is to apply the idea of overlapping batches to MRP.  
The motivation behind using overlapping in this context is twofolds. 
First, overlapping could help to obtain better variance estimators with simple data reuse, resulting in improved CIs.
%Especially for problems where sampling error is dominant, this simple data reuse method could yield better estimators.
Second, in some cases, generating the realizations $\xit^1, \xit^2, \dots$ to form sampling approximations of (\ref{eq:sto_prog}) can be computationally expensive and overlapping makes the most of the generated observations.

%The following are four specific contributions of this article. 
The contributions of this article are as follows. 
\begin{itemize}
\item[(i)] We formulate an overlapping batch means variant of MRP, where the degree of overlap can be chosen by the user. 
%This paper applies OBM to assessing solution quality in stochastic optimization. 
%Specifically, an overlapping batch means variant of MRP is formulated, where the degree of overlap can be chosen by the user. 
This application of OBM is novel because each batch mean estimator includes an internal minimization, which has not been studied in the literature.
%\vspace*{-0.1in}

@@

\item[(iii)] We  present computational results to examine small-sample behavior. 
Our experiments indicate that the use of overlapping estimators results in gap estimators achieving lower variance than their nonoverlapping counterparts, along with comparable bias and interval coverage---for both small- and large-sample cases.

\item[(iv)] As the degree of overlap increases, the number of batches and optimization problems solved increases because we need to solve a sampling problem per batch. 
At the same time, variances of the estimators are lowered, and when warmstarting is available, solution time per batch can decrease, resulting in efficiency gains. 
We explore this trade-off between computation time and lower variances.
% and efficiency gains. 
%We explore the overall computation time versus lower variances and efficiency gains trade-off. 
Our results indicate that when warmstarting is effective, overlapping is computationally more efficient relative to estimators formed via nonoverlapping batches.
%Our results indicate that overlapping can indeed increase computational efficiency, especially when warmstarting is effective. 
We provide guidelines on selecting the degree of overlap based on computational efficiency and demonstrate efficiency on a newsvendor problem. 
\end{itemize}

An earlier version of the paper appeared in \citep{love2011overlapping}. 
This paper contains a number of additional results. 
% compared to its earlier conference proceedings version. 
Most of the theoretical results are new, including asymptotically lower variances of the overlapping estimators, asymptotic validity of the interval estimators, and the {\it almost sure} convergence of the variance estimator. 
These results are achieved through an assumption on the fast rate of convergence of the sampling problem's solution to the unique optimal solution, which is satisfied by several classes of (\ref{eq:sto_prog}); see \S \ref{subsec:assumptions}.
We also present additional computational results, including those from a large-scale problem, and study computational efficiency of the overlapping MRP method.

@@

%In this paper, we extend MRP by overlapping the batches similar to the overlapping batch means (OBM) introduced by Meketon and Schmeiser \citep{Meketon1984}.

%We explore the use of overlapping batches for assessing solution quality in stochastic programs.
%Similar to Welch \citep{Welch1987} and Song and Schmeiser \citep{Song1992}, we consider partial overlap between batches to reduce computational effort in our setting.
%The contributions of this paper are:
%\begin{itemize}
%	\item Formulate the overlapping multiple replications procedure for stochastic programs.
%	\item Give conditions under which the point estimators are strongly consistent (that is, they converge almost surely (a.s.) as opposed in to in probability; referred to simply as consistent from this point on).
%	\item Demonstrate that the classical variance reduction results are achieved for stochastic programs.
%	\item Show the asymptotic validity of the resulting confidence intervals.
%	\item Present computational results on two-stage stochastic linear and integer programs with recourse to examine small sample behavior.  
%		Our experiments indicate that the asymptotic variance reduction is achieved with small sample sizes, while bias and coverage probability are unaffected.
%\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%   %    %    %    %    %    %

%\bigskip

% We present variants of . .

%\bigskip

%Batch means have been used in stochastic programming to assess solution quality and we show that. . . 
%\bigskip

%What happens when overlapping batches are used in a stochastic optimization setting?  Can we obtain the same benefits from overlapping batches as we do in a non-optimized setting?  In this paper, we answer the second question in the affirmative. We provide conditions under which the same benefits can be obtained by overlapping for a class of stochastic optimization problems. A natural setting where batch means is used in stochastic optimization is solution validation. This is referred to as the Multiple Replication Procedure, where each replication is essentially a batch means.

%\bigskip 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%   %    %    %    %    %    %

@@

Overlapping batch means reuses a subset of the observations from one batch to form the next batch. 
As an example, suppose each batch contains $m$ observations and that there are a total of $n=2m$ observations.
In the nonoverlapping case, there are only two batches:
% the first batch contains observations $y^1, y^2, \dots, y^m$ and the second batch contains observations $y^{m+1}, y^{m+2}, \dots, y^{2m}$.  
$y^1, y^2, \dots, y^m$ and $y^{m+1}, y^{m+2}, \dots, y^{2m}$.  
In the overlapping batches of \citet{Meketon1984}, the first batch is as before, the second batch now consists of $y^2, y^3, \dots, y^{m+1}$, the third batch $y^3, y^4, \dots, y^{m+2}$, and so on.  
This is what we call the \emph{maximally-overlapping} batch means method, because $m-1$ observations are common to---and only one observation changes between---adjacent batches.  
It is possible to vary the degree of overlap between neighboring batches. 
This generalization was first studied by \citet{Welch1987}, who also established the relationship between OBM and spectral estimators; see also \citep{Song1992}.
%
The overlapping batch means 
%formed using the batch nonoverlap parameter $1 \leq \gamma < m$ 
are no longer independent. 
However, by simply reusing the data in this fashion one obtains a better variance estimator. 
That is, the overlapping variance estimator has (i) nearly the same bias as the standard nonoverlapping variance estimator, and it has (ii)
%the overlapping estimator has 
a lower asymptotic variance than its nonoverlapping counterpart.

In this paper, we use the {\it batch nonoverlap parameter} $1 \leq \gamma \leq m$ to denote how much neighboring batches do {\it not} overlap.  
For instance, $\gamma = m$ corresponds to the case of nonoverlapping batches and $\gamma = 1$ corresponds to the maximally-overlapping case of \citet{Meketon1984}.  
In general, $\gamma$ can take any integer value between and including $[1,m]$. 
Let $\lfloor \cdot \rfloor$ be the greatest integer that is less than or equal to its argument. 
The sample mean of each batch is calculated similarly, 
$
\yb_j = \frac{1}{m} \sum_{i=1}^m y^{\gamma(j-1) + i},\ \  j = 1, 2, \dots, \nb, 
$
now taking into account $\gamma$.  
Here, $\nb$ batches are formed given $n$, $m$, and $\gamma$.  
The variance estimator is updated to
\begin{equation} \label{eq:variance_gamma}
	\vartg{\ybb} = \frac{1}{\left( \tfrac{n}{m} - 1 \right) \left( \nb \right)}  \sum_{j=1}^{\nb} (\yb_j - \ybb)^2.
\end{equation}
With $\gamma=m$ and $n=mk$, (\ref{eq:variance_gamma}) reduces to (\ref{eq:var}).
With $\gamma=1$ (the maximally-overlapping case), the denominator in (\ref{eq:variance_gamma}) is slightly different than the one in  \citep{Meketon1984}.
This slightly different denominator 
makes $\vto{\ybb}$ an unbiased estimator for i.i.d.\ data for all finite $m$ and $n$ with the same asymptotic benefits of nearly the same bias and lower variances relative to its counterpart obtained via nonoverlapping batches \citep{Song1992}.

The amount of asymptotic reduction in $\var{\vartg{\ybb}}$ for a given $0\leq \gamma< m$ relative its nonoverlapping counterpart 
%$\var{\vtm{\ybb}}$ 
depends on the asymptotic ratio of $\gamma/m$, which we denote by $\gammab$. 
The maximally-overlapping case ($\gamma = 1$) corresponds to $\gammab=0$, and the nonoverlapping case corresponds to $\gammab=1$. 
\cite{damerdji1994strong,damerdji1995mean} studied batching structures that guarantee almost sure and mean-square consistency of $\vartg{\ybb}$ in the nonoverlapping and maximally-overlapping cases. 
Under appropriate conditions, the results of \cite{damerdji1995mean}, along with those of \cite{Meketon1984} and \cite{Welch1987}, establish that 
\begin{equation}
\label{eq:varvarred}
\frac{n}{m}\var{n \vartg{\ybb}} \rightarrow \beta_{\gammab} \sigma^4
\end{equation}
as the batch size $m$ and then the number of batches $n/m$ tend to infinity. 
In (\ref{eq:varvarred}), $\beta_1 = 2$ for nonoverlapping batches, $\beta_0 = \frac{4}{3}$ for maximally overlapping batches, and more generally, $\beta_{\gammab}$ is twice the value of 
\begin{equation} \label{eq:var_reduct_formula}
	\gammab \left( 1 + 2 \sum_{j=1}^{\cogb-1} (1-j\gammab)^2 \right),
\end{equation}
where $\lceil \cdot \rceil$ denotes the smallest integer that is greater than or equal to its argument.
If $\ogb$ is an integer, (\ref{eq:var_reduct_formula}) simplifies to $\frac{2+\gammab^2}{3}$.
This means that the maximally-overlapping variance estimator has a variance that is two-thirds ($66.67\%$) of the nonoverlapping estimator.
As the degree of overlap decreases, the variance increases. 
When only half of the observations overlap ($\gammab = 1/2$ or $\gamma=m/2$), for instance, the variance is 75\% of the nonoverlapping counterpart.

A $(1-\alpha)$-level approximate CI on the mean can be formed using overlapping batches by $\ybb \pm t_{d_{\gammab}(k-1),\alpha/2}$ $\sqrt{\vartg{\ybb}}$, where the degrees of freedom increase $d_{\gammab}$ is the multiplicative inverse of (\ref{eq:var_reduct_formula}) \citep{Welch1987}.  
%If $\ogb$ is an integer, $d_{\gammab} = \frac{3}{2+\gammab^2}$.  
For the maximally-overlapping batch means, the degrees of freedom is $d_{\gammab} = \frac{3}{2}$.

@@

%Given a candidate solution $\xh \in \X$ to (\ref{eq:sto_prog}), the task is to estimate its optimality gap $\e{f(\xh,\xit)} - \zs$.  
%Recall that MRP uses the upper bound on the optimality gap, $\e{f(\xh,\xit)} - \e{\zs_m}$ (for a given sample size $m$), to construct a point estimator and a CI using the nonoverlapping batches method described above.

Taking $n$ observations of the random variables $\xit$ and splitting them into $k$ {\it nonoverlapping} batches of size $m$, define 
\begin{align} 
	\gb_j  & = \frac{1}{m} \sum_{i = 1}^{m} f(\xh, \xit^{m(j-1)+i}) - \min_{\x \in X} \frac{1}{m} \sum_{i=1}^{m} f(\x, \xit^{m(j-1)+i}), \ \ \ \ j = 1,2, \dots, k \label{eq:mrp_gap}\\
         & = \frac{1}{m} \sum_{i = 1}^{m} \left[ f(\xh, \xit^{m(j-1)+i}) -  f(\xs_j, \xit^{m(j-1)+i}) \right], \ \ \ \ j = 1,2, \dots, k, \notag
\end{align}
where $\xs_j$ denotes an optimal solution to (\ref{eq:sto_prog_m}) of batch $j$ given on the right-hand side of (\ref{eq:mrp_gap}).    
Since we are assessing the quality of a given solution $\xh \in \X$, we suppress it from the notation $\gb_j$.
The optimality gap estimator $\gb_j$ is similar to the nonoverlapping $\yb_j$ with individual observations $y^{m(j-1)+i} = f(\xh,\xit^{m(j-1)+i}) - f(\xs_j,\xit^{m(j-1)+i})$, for $i=1,\ldots,m$ and $j=1,\ldots, k$.
Notice that in this case, $y^i$ not only depends on observation $i$ of batch $j$, $\xit^{m(j-1)+i}$, but all the observations in that batch $\xit^{m(j-1)+1},\xit^{m(j-1)+2},\ldots,\xit^{m(j-1)+m}$ because $\xs_j$ is the optimal solution to (\ref{eq:sto_prog_m}) given in (\ref{eq:mrp_gap}).   
As before, after $k$ estimates are obtained, the overall mean of these estimates $\gbb = \frac{1}{k} \sum_{j=1}^k \gb_j$ provides a point estimator of the optimality gap.  
The variance estimator is obtained as in (\ref{eq:var}) by $\varh{\gbb} = \frac{1}{k} \frac{1}{k-1} \sum_{j=1}^k (\gb_j - \gbb)^2$, which results in an approximate $(1-\alpha)$-level one-sided CI $\left[0, \gbb + t_{k-1,\alpha} \sqrt{\varh{\gbb}} \right]$.

@@

Another difference between our setting and simulation output analysis is that once the data are generated through a simulation, they can be reused without much additional computational effort to obtain the OBM variance estimator.  
In this setting, the computational effort may increase with data reuse because we need to solve a sampling problem (SP$_m$) for each batch.  
The increase in computation time can be alleviated in two ways: (i) by increasing $\gamma$ to partially overlap the batches, resulting in a smaller number of batches, and (ii) using warmstarting to quickly solve the sampling approximations with overlapping samples. 
%We use the batch nonoverlap parameter $\gamma$ as before to allow for varying degrees of overlap in OMRP. 
We examine the effect of warmstarting and any resulting computational efficiency gains in \S \ref{ssec:compeff}. 
We are now ready to define the OMRP estimators.

%First, near-optimal variance reduction can be obtained by partially overlapping the batches \citep{Welch1987,Song1992}.  
%Partial overlapping results in a smaller number of batches; hence, fewer optimization problems need to be solved.
%Second, in many solution methods, warmstarting can be used to quickly solve the sampling approximations with overlapping samples, considerably reducing solution time.
%In this paper, we consider a parametrized degree of overlap, where the degree of overlap is at the user's discretion, and examine this both theoretically and empirically. 
% THIS LAST SENTENCE NEEDS TO CHANGE!

@@

The results on overlapping batch means discussed in \S \ref{ssec:overlap} occur in the limit as $n, m, k=n/m \rightarrow \infty$.  
Let $n_l, m_l$, and $k_l$ be sequences of numbers satisfying these requirements, then the limits will be taken as $l \rightarrow \infty$.  
We assume a batching structure where $m_l = \lfloor n_l ^r \rfloor$ for some $0<r<1$, as is typical in simulation, and $n_l$ tends to infinity as $l \rightarrow \infty$.  
In addition, we may desire that the batch nonoverlap parameter change with $l$, to ensure that $\gamma_l / m_l = \gammab_l$ converges to the constant $\gammab$.  
The OMRP estimators are defined as follows:
\begin{align}
	\gb_{lj} & = \frac{1}{m_l} \sum_{i=1}^{m_l} \left[ f(\xh,\xit^{\gamma_l(j-1)+i}) - f(\xs_j,\xit^{\gamma_l(j-1)+1}) \right],\ \ \ j = 1, 2, \dots, \nbl, \label{eq:gbar} \\
	\gbb_l & = \frac{1}{n_l} \sum_{i=1}^{n_l} \frac{1}{|B(i)|} \sum_{j \in B(i)} \left[ f(\xh,\xiti) - f(\xs_j,\xiti) \right], \label{eq:gbb} \\
	VG_l & = \dfrac{1}{\left( \tfrac{n_l}{m_l} - 1 \right) \left(\nbl\right)} \sum_{j=1}^{\nbl} (\gb_{lj} - \gbb_l)^2. \label{eq:vg}
\end{align}

@@

In this section, we provide conditions under which several asymptotic properties hold for OMRP estimators. 
First, we show the consistency of the point estimators of optimality gap, $\gbb_l$, and variance, $n_l VG_l$. 
Then, we provide conditions under which the asymptotically lower variances are observed for the OMRP variance estimator relative to nonoverlapping MRP estimator.  
Finally, we show the asymptotic validity of the OMRP interval estimator.

>>>>

The Overlapping Batch Means (OBM) method, proposed by \citet{Meketon1984}, is commonly used in simulation to obtain variance estimators having variances that are comparatively lower than their nonoverlapped counterparts. 
In the context of steady-state simulation, OBM is often used to estimate the variance of the sample mean, which is itself an estimator for the mean. 
\cite{SAH90} use overlapping estimators to estimate the variance of estimators for non-means. 
More recently, overlapping estimators based on standardized time series (as opposed to batch means) have been evaluated \citep{Alexopoulos01012007,Alexopoulos2007}; see also \citep{Meterelliyoz_etal_12}.
To the best of our knowledge, OBM has never before been studied in a stochastic optimization context. 
In this paper, we show that the same benefits of nearly the same bias and lower variances relative to their nonoverlapped counterparts can be obtained from overlapping batches when used in a stochastic optimization setting.

@@

The optimality gap $\e{f(\xh,\xit)} - \zs$ often cannot be calculated exactly. First, for a given candidate solution $\xh \in X$, evaluation of $\e{f(\xh,\xit)}$ typically involves a difficult multidimensional integral. 
Second, we do not know $\zs$. 
In many fields of optimization, the second problem is alleviated by obtaining lower bounds on $\zs$ through relaxations such as integrality, Lagrangian, or semidefinite relaxations.
As a result, it is common to evaluate an upper bound on the optimality gap. 
This motivates estimating the upper bound on the optimality gap in stochastic programming by using Monte Carlo simulation.
Given a sample size $m$, an upper bound on the optimality gap can be obtained by $$
\e{f(\xh,\xit)} - \zs \leq \e{f(\xh,\xit)} - \e{\zs_m}
$$ 
due to the inequality $\e{\zs_m} \leq \zs$ \citep{Mak1999,norkin_pflug_ruszczynski_98}.
This bound improves as the sample size increases; that is, $\e{\zs_m} \leq \e{\zs_{m+1}} \leq \zs$.  
A straightforward estimate of $\e{f(\xh,\xit)}$ is the sample mean, $\frac{1}{m} \sum_{i=1}^m f(\xh,\xiti)$. 
Instead of $\e{\zs_m}$, we can simply use $\zs_m$.  
The resulting point estimator of the optimality gap is then 
$$
\frac{1}{m} \sum_{i=1}^m f(\xh,\xiti) - \zs_m.
$$ 
Here, we assume the same observations $\xit^1, \dots, \xit^m$ are used in both terms.

Computing the above optimality gap estimate involves solving an optimization problem (\ref{eq:sto_prog_m}) to obtain a lower bound estimator of $\zs$, $\zs_m$, which complicates the statistical analysis. 
As mentioned above, to enable statistical inference, the MRP of \citet{Mak1999} generates $k$ independent estimators, each using sample size $m$ ($k$ nonoverlapping batches of size $m$), and averages them to obtain a point estimator.  
The sample variance of these estimators is used to form confidence intervals (CIs) for the optimality gap (we further review MRP in \S \ref{ssec:mrp}).  
An advantage of MRP is its applicability to a wide range of problems.  
With i.i.d.\ sampling, $f$ can be linear or nonlinear, $\X$ can include integrality constraints or not.  
It is also easy to implement; thus, MRP has been applied to a wide variety of problems, including finance, supply chain network design, etc.; see, e.g., \citep{bertocchi_etal_99,janjarassuk_linderoth_08,santoso_ahmed_etal_05}.  
Recently, the approach of (nonoverlapping) batching has been used for assessing solution quality of stochastic programs with finitely many expected value \citep{wang_ahmed_08} and stochastic dominance \citep{hu2010sample} constraints.

@@

Our aim is to apply the idea of overlapping batches to MRP.  
The motivation behind using overlapping batches in this context is twofold. 
First, overlapping could help to obtain better variance estimators with simple data reuse, resulting in improved CIs.
%Especially for problems where sampling error is dominant, this simple data reuse method could yield better estimators.
Second, in some cases, generating the realizations $\xit^1, \xit^2, \dots$ to form sampling approximations of (\ref{eq:sto_prog}) can be computationally expensive, and in these cases, overlapping makes the most of the generated observations by reusing data.

%The following are four specific contributions of this article. 
The contributions of this article are as follows: 
\begin{itemize}
\item[(i)] We formulate an overlapping batch means variant of MRP, where the degree of overlap can be chosen by the user.  
This application of OBM is novel because each batch mean estimator includes an internal minimization, which has not been studied in the literature.
%\vspace*{-0.1in}

@@

\item[(iii)] We examine small-sample behavior of the estimators through numerical examples. 
Our experiments indicate that the use of overlapping estimators results in gap estimators achieving lower variance than their nonoverlapping counterparts, along with comparable bias and interval coverage---for both small- and large-sample cases.

\item[(iv)] As the degree of overlap increases, the number of batches---and optimization problems solved---increases because we need to solve a sampling problem per batch. 
At the same time, variances are lowered, and when warm starting is available, solution time per batch can decrease, yielding efficiency gains. 
We explore this trade-off between computation time and lower variances.
Our results indicate that when warm starting is effective, overlapping estimators are computationally more efficient than those formed via nonoverlapping batches.
Based on this study, we provide guidelines on selecting the degree of overlap and demonstrate efficiency on a newsvendor problem. 
\end{itemize}

An earlier version of the paper appeared in \citep{love2011overlapping}. 
Most of the theoretical results in the current version of the paper are new, including {\it almost sure} convergence and asymptotically lower variances of the overlapping variance estimators, and asymptotic validity of the interval estimators. 
These results are achieved through a stronger assumption on the rate of convergence of the sampling problem's solution to the unique optimal solution, which is satisfied by several classes of (\ref{eq:sto_prog}); see \S \ref{subsec:assumptions}.
We also present additional computational results, including those from a large-scale problem, and study computational efficiency of the overlapping MRP method.

@@

@@

Overlapping batch means modifies nonoverlapping batch means by reusing a subset of observations from one batch to form the next batch. 
%As an example, suppose each batch contains $m$ observations and that there are a total of $n=2m$ observations.
%In the nonoverlapping case, there are only two batches:
%% the first batch contains observations $y^1, y^2, \dots, y^m$ and the second batch contains observations $y^{m+1}, y^{m+2}, \dots, y^{2m}$.  
%$y^1, y^2, \dots, y^m$ and $y^{m+1}, y^{m+2}, \dots, y^{2m}$.  
%In the overlapping batches of \citet{Meketon1984}, the first batch is as before, the second batch now consists of $y^2, y^3, \dots, y^{m+1}$, the third batch $y^3, y^4, \dots, y^{m+2}$, and so on.  
%This is what we call the \emph{maximally overlapping} batch means method, because $m-1$ observations are common to---and only one observation changes between---adjacent batches.  
%It is possible to vary the degree of overlap between neighboring batches. 
%This generalization was first studied by \citet{Welch1987}, who also established the relationship between OBM and spectral estimators; see also \citep{Song1992}.
%%
In this paper, we use the {\it batch nonoverlap parameter} $1 \leq \gamma \leq m$ to denote how much neighboring batches do {\it not} overlap.  
For instance, $\gamma = m$ corresponds to the case of nonoverlapping batches and $\gamma = 1$ corresponds to the maximally overlapping case of \citet{Meketon1984}.  
In general, $\gamma$ can take any integer value between and including $[1,m]$. 
Let $\lfloor \cdot \rfloor$ be the greatest integer that is less than or equal to its argument. 
The sample mean of each batch is calculated similarly, 
$
\yb_j = \frac{1}{m} \sum_{i=1}^m y^{\gamma(j-1) + i},\ \  j = 1, 2, \dots, \nb, 
$
now taking into account $\gamma$.  
Here, $\nb$ batches are formed given $n$, $m$, and $\gamma$.  
The variance estimator is updated to
\begin{equation} \label{eq:variance_gamma}
	\vartg{\ybb} = \frac{1}{\left( \tfrac{n}{m} - 1 \right) \left( \nb \right)}  \sum_{j=1}^{\nb} (\yb_j - \ybb)^2.
\end{equation}

The overlapping batch means 
formed using $1 \leq \gamma < m$ 
are no longer independent. 
However, by simply reusing the data in this fashion one obtains a better variance estimator. 
That is, the overlapping variance estimator has (i) nearly the same bias as the standard nonoverlapping variance estimator, and it has (ii)
%the overlapping estimator has 
a lower asymptotic variance than its nonoverlapping counterpart. 
Observe that with $\gamma=m$ and $n=mk$, (\ref{eq:variance_gamma}) reduces to (\ref{eq:var}).
With $\gamma=1$ (the maximally overlapping case), the denominator in (\ref{eq:variance_gamma}) is slightly different than the one in  \citep{Meketon1984}.
This slightly different denominator 
makes $\vto{\ybb}$ an unbiased estimator for i.i.d.\ data for all finite $m$ and $n$ with the same asymptotic benefits of nearly the same bias and lower variance relative to its counterpart obtained via nonoverlapping batches \citep{Song1992}.

The amount of asymptotic reduction in $\var{\vartg{\ybb}}$ for a given $1\leq \gamma< m$ relative its nonoverlapping counterpart 
%$\var{\vtm{\ybb}}$ 
depends on the asymptotic ratio of $\gamma/m$, which we denote by $\gammab$. 
The maximally overlapping case ($\gamma = 1$) corresponds to $\gammab=0$, and the nonoverlapping case corresponds to $\gammab=1$. 
\cite{damerdji1994strong,damerdji1995mean} studied batching structures that guarantee almost sure and mean-square consistency of $\vartg{\ybb}$ in the nonoverlapping and maximally overlapping cases. 
Under appropriate conditions, the results of \cite{damerdji1995mean}, along with those of \cite{Meketon1984} and \cite{Welch1987}, establish that 
\begin{equation}
\label{eq:varvarred}
\frac{n}{m}\var{n \vartg{\ybb}} \rightarrow \beta_{\gammab} \sigma^4
\end{equation}
as the batch size $m$ and then the number of batches $n/m$ tend to infinity. 
In (\ref{eq:varvarred}), $\beta_1 = 2$ for nonoverlapping batches and $\beta_0 = \frac{4}{3}$ for maximally overlapping batches.  
More generally, $\beta_{\gammab}$ is twice the value of 
\begin{equation} \label{eq:var_reduct_formula}
	\gammab \left( 1 + 2 \sum_{j=1}^{\cogb-1} (1-j\gammab)^2 \right),
\end{equation}
where $\lceil \cdot \rceil$ denotes the smallest integer that is greater than or equal to its argument \citep{Welch1987}.
If $\ogb$ is an integer, (\ref{eq:var_reduct_formula}) simplifies to $\frac{2+\gammab^2}{3}$.
This means that the maximally overlapping variance estimator has a variance that is two-thirds ($66.67\%$) of the nonoverlapping estimator.
As the degree of overlap decreases, the variance increases. 
When only half of the observations overlap 
($\gammab = 1/2$),
%($\gammab = 1/2$ or $\gamma=m/2$),
for instance, the variance is 75\% of the nonoverlapping counterpart.

A $(1-\alpha)$-level approximate CI on the mean can be formed using overlapping batches by $\ybb \pm t_{d_{\gammab}(k-1),\alpha/2}$ $\sqrt{\vartg{\ybb}}$, where the degrees of freedom increase $d_{\gammab}$ is the multiplicative inverse of (\ref{eq:var_reduct_formula}). %\citep{Welch1987}.  
%If $\ogb$ is an integer, $d_{\gammab} = \frac{3}{2+\gammab^2}$.  
For the maximally overlapping batch means, the degrees of freedom is $d_{\gammab} = \frac{3}{2}$.

@@

Taking $n$ observations of $\xit$ and splitting them into $k$ {\it nonoverlapping} batches of size $m$, define
\begin{align} 
	\gb_j  & = \frac{1}{m} \sum_{i = 1}^{m} f(\xh, \xit^{m(j-1)+i}) - \min_{\x \in X} \frac{1}{m} \sum_{i=1}^{m} f(\x, \xit^{m(j-1)+i}), \ \ \ \ j = 1,2, \dots, k \label{eq:mrp_gap}\\
         & = \frac{1}{m} \sum_{i = 1}^{m} \left[ f(\xh, \xit^{m(j-1)+i}) -  f(\xs_j, \xit^{m(j-1)+i}) \right], \ \ \ \ j = 1,2, \dots, k, \notag
\end{align}
where $\xs_j$ denotes an optimal solution to (\ref{eq:sto_prog_m}) of batch $j$ given on the right-hand side of (\ref{eq:mrp_gap}).    
Since we are assessing the quality of a given solution $\xh \in \X$, we suppress it from the notation $\gb_j$.
The optimality gap estimator $\gb_j$ is similar to the nonoverlapping $\yb_j$ with individual observations $y^{m(j-1)+i} = f(\xh,\xit^{m(j-1)+i}) - f(\xs_j,\xit^{m(j-1)+i})$, for $i=1,\ldots,m$ and $j=1,\ldots, k$.
Notice that in this case, $y^i$ not only depends on observation $i$ of batch $j$, $\xit^{m(j-1)+i}$, but all the observations in that batch $\xit^{m(j-1)+1},\xit^{m(j-1)+2},\ldots,\xit^{m(j-1)+m}$ because $\xs_j$ is an optimal solution to (\ref{eq:sto_prog_m}) given in (\ref{eq:mrp_gap}).   
As before, after $k$ batch-means estimators are obtained, the overall mean of these estimators $\gbb = \frac{1}{k} \sum_{j=1}^k \gb_j$ provides a point estimator of the optimality gap.  
The variance estimator is obtained as in (\ref{eq:var}) by $\varh{\gbb} = \frac{1}{k} \frac{1}{k-1} \sum_{j=1}^k (\gb_j - \gbb)^2$, which results in an approximate $(1-\alpha)$-level one-sided CI $\left[0, \gbb + t_{k-1,\alpha} \sqrt{\varh{\gbb}} \right]$.

@@

Another difference between our setting and simulation output analysis is that once the data are generated through a simulation, they can be reused without much additional computational effort to obtain the OBM variance estimator.  
In this setting, the computational effort may increase with data reuse because we need to solve a sampling problem (SP$_m$) for each batch.  
The increase in computation time can be alleviated in two ways: (i) by increasing $\gamma$ to partially overlap the batches, resulting in a smaller number of batches (but a higher variance), and (ii) using warm starting to quickly solve the sampling approximations with overlapping observations. 
%We use the batch nonoverlap parameter $\gamma$ as before to allow for varying degrees of overlap in OMRP. 
We investigate the balance of computational efficiency gains from warm starting versus variability of the estimators in \S \ref{ssec:compeff}. 
We are now ready to define the OMRP estimators.

@@

The results on overlapping batch means discussed in \S \ref{ssec:overlap} occur in the limit as $n, m, k=n/m \rightarrow \infty$.  
Let $n_l, m_l$, and $k_l$ be sequences of numbers satisfying these requirements, then the limits will be taken as $l \rightarrow \infty$.  
We assume a batching structure where $m_l = \lfloor n_l ^r \rfloor$ for some $0<r<1$, as is typical in simulation, and $n_l$ tends to infinity as $l \rightarrow \infty$.  
In addition, we may desire that the batch nonoverlap parameter change with $l$, to ensure that $\gamma_l / m_l = \gammab_l$ converges to the constant $\gammab$.  
The OMRP estimators are defined as follows:
\begin{align}
	\gb_{lj} & = \frac{1}{m_l} \sum_{i=1}^{m_l} \left[ f(\xh,\xit^{\gamma_l(j-1)+i}) - f(\xs_j,\xit^{\gamma_l(j-1)+1}) \right],\ \ \ j = 1, 2, \dots, \nbl, \label{eq:gbar} 
\end{align}
\begin{align}
	\gbb_l & = \frac{1}{n_l} \sum_{i=1}^{n_l} \frac{1}{|B(i)|} \sum_{j \in B(i)} \left[ f(\xh,\xiti) - f(\xs_j,\xiti) \right], \label{eq:gbb} \\
	VG_l & = \dfrac{1}{\left( \tfrac{n_l}{m_l} - 1 \right) \left(\nbl\right)} \sum_{j=1}^{\nbl} (\gb_{lj} - \gbb_l)^2. \label{eq:vg}
\end{align}

@@

In this section, we provide conditions under which several asymptotic properties hold for OMRP estimators. 
First, we show the consistency of the point estimators of optimality gap, $\gbb_l$, and variance, $n_l VG_l$. 
Then, we provide conditions under which the OMRP variance estimator achieves asymptotically lower variances relative to the MRP estimator.  
Finally, we show the asymptotic validity of the OMRP interval estimator.
